|

### 2020-02-03 paper/cogsci/paper_initial_submission_blinded.Rmd
cleaned paper repo, added data links, last typos, rendered pdfs

author-information: > 
    \author{{\large \bf Bria Long (bria@stanford.edu)} 
    \AND {\large \bf George Kachergis (kachergis@stanford.edu)} 
    \AND {\large \bf Ketan Jay Agarwal (agrawalk@stanford.edu)} 
    \AND {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ 
    Department of Psychology, Street Address \\ Stanford, CA 91305 USA}

abstract: >
        The faces and hands of caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of nearly 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validated the use of a modern convolutional neural network for pose detection (OpenPose) for the detection of faces and hands. We then applied this model to the entire dataset, and found a higher proportion of hands in view than previous reported and a moderate decrease the proportion of faces in children's view across age. In addition, we found variability in the proportion of faces/hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience.

@@

Beyond these early months, infants' motor and cognitive abilities mature, leading to a vastly different perspectives on the world. For example, crawlers see fewer faces and hands than do walking children [@sanchez2018postural; @franchak2017see; @kretch2014] as well as different views of objects [@smith2011not]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, and their perspective starts to capture views of hands manipulating objects [@Fausey2016]. In turn, caregivers may also start to use their hands with more communicative intent, directing infants' attention by pointing and gesturing to different events and objects during play [@yu2013joint].

Here, we examine the social information present in the infant visual perspective---the presence of faces and hands---by analyzing a longitudinal collection of nearly 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCAM dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYcam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008; @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

@@

## Detection Method
To automatically annotate the millions of frames in SAYcam, we used OpenPose [@Cao2018openpose; @Simon2017hand]. This pose detector ^[https://github.com/CMU-Perceptual-Computing-Lab/openpose] provided the locations of 18 body parts (ears, nose, wrists, etc.). A convolutional neural network was used for initial anatomical detection, and part affinity fields were subsequently applied for part association to produce a series of body part candidates. Once these body part candidates were matched to a single individual in the frame, they were finally assembled into a pose. Thus, while we only made use of the outputs of the face and hand detections, the entire set of pose information from an individual was used to determine the presence of a face/hand, making the process much more robust to occlusion than methods optimized to detect only faces or hands. Note, however, that these face/hand detections are reliant on the detection of at least a partial pose, so some very up-close views of faces/hands may go undetected.

@@

As has been observed in other studies on automated annotation of headcam data [e.g. @frank2013; @sanchez2018postural; @bambach2015lending], detection tasks that are easy in third-person video can be quite challenging in egocentric videos, due to difficult angles and sizes as well as substantial occlusion. For example, the infant perspective often contains non-canonical viewpoints of faces (e.g., looking up at a caregiver's chin) as well as partially-occluded or oblique viewpoints of both faces and hands. Furthermore, hand detection tends to be a harder problem than face detection [@Simon2017hand; @bambach2015lending]. We thus expected overall performance to be lower in these naturalistic videos than on either photos taken from the adult perspective or in egocentric videos in controlled, laboratory settings [e.g., @sanchez2018postural].

@@

```{r FacesAndHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3.5, fig.cap = "Proportion of faces and hands seen as a function of age for each child in the dataset. Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range. Dashed lines show estimated trend lines from proportion of faces/hands in view when detections are restricted to the upper 60 percent of the FOV, reducing the contribution of children's own hands; dotted lines show estimated trend lines for human annotations on the 24K gold set."}

@@

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child. Data from videos were binned according to the age of the child (in weeks).  First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:FacesAndHands}); in contrast, we did not observe an increase in the proportion of hands in view, but rather a slight decrease. These effects were quantified with two separate linear mixed-models (see Tables 1 & 2).^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.]

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Figure \ref{fig:FacesAndHands}) or OpenPose annotations on the entire dataset (see solid lines in Figure \ref{fig:FacesAndHands}). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. Nonetheless, one reason this could be the case is the much larger field of view that was captured by the cameras used in this study: These cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible, leading to a larger field of view (109 degrees horizontal x 70 degrees vertical) than in many previous studies; for example, in @Fausey2016 the FOV was 69 x 41 degrees. This larger FOV may have allowed the SAYcam cameras to capture not only the presence of a social partner’s hands interacting with objects or gestures, but also the children’s own hands, leading to more frequent hand detections.

```{r}
## replace NAs for computing prop hands in cropped anlaysis
d_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) 
```

As children's hands tended to occur in the lower visual field (see Figure \ref{fig:handLocation}), we thus re-analyzed the entire dataset while restricting our analysis to the upper 60% of the field of view, decreasing the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)*100`% to `r round(mean(d_cropped$hand_detected),2)*100`%, but only decreased face detections from `r round(mean(d_all$face_openpose),2)*100`% to `r round(mean(d_cropped$face_detected),3)*100`%. This cropping likely removed both the majority of detections of children's own hands but does also remove some detections of adult hands (see Figure \ref{fig:handLocation}), especially as OpenPose was biased to miss children's hands when they were in view. Nonetheless, within this modified field of view, we still observed more hand detections than face detections (see dashed lines in Figure \ref{fig:FacesAndHands}).

@@

Here, we analyzed the social information in view in a dense, longitudinal dataset, applying a modern computer-vision model to quantify the proportion of hands and faces seen from each of three children's egocentric perspective from 6 to 32 months of age. This analysis has yielded a better understanding of infants' evolving access to social information. We found a moderate decrease across age in the proportion of faces in view in the videos, in keeping with previous work [@Fausey2016]. However, this finding is particularly notable given that, in previous cross-sectional data, this effect seems to be most strongly driven by infants younger than 4 months of age [e.g., @Fausey2016; @Jayaraman2015; @Sugden2014] who see both more frequent and more persistent faces [@Jayaraman2018].

@@

A coarse analysis based on the location the videos were filmed in further highlights the variability of the social information in view during different activites, showing differences across locations and between individual children. Within a given, well-defined context---e.g., mealtime in kitchens---S saw more faces than A, and S saw more faces in the kitchen than in other locations. This variability likely stems from the fact that there are at least three ways to feed a young child: 1) sitting in front of the child, facing them as they sit in a high chair; 2) sitting behind the child, holding them as they face outward, and 3) sitting side by side. Each of these positions offer the child differing degrees of visual access to face and hands. While the social information in view may be variable across children in different activitiy contexts, these analyses suggest they could be stable within a given child's day-to-day experience.

>>>>

author-information: >

abstract: >
        The faces and hands of caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of nearly 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validated the use of a modern convolutional neural network of pose detection (OpenPose) for the detection of faces and hands. We then applied this model to the entire dataset, and found a higher proportion of hands in view than previous reported and a moderate decrease the proportion of faces in children's view across age. In addition, we found variability in the proportion of faces/hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience.

@@

Beyond these early months, infants' motor and cognitive abilities mature, leading to vastly different perspectives on the world. For example, crawlers see fewer faces and hands than do walking children [@sanchez2018postural; @franchak2017see; @kretch2014] as well as different views of objects [@smith2011not]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, and their perspective starts to capture views of hands manipulating objects [@Fausey2016]. In turn, caregivers may also start to use their hands with more communicative intent, directing infants' attention by pointing and gesturing to different events and objects during play [@yu2013joint].

Here, we examine the social information present in the infant visual perspective---the presence of faces and hands---by analyzing a longitudinal collection of nearly 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYCam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008; @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

@@

## Detection Method
To automatically annotate the millions of frames in SAYCam, we used a pose detector, OpenPose ^[https://github.com/CMU-Perceptual-Computing-Lab/openpose]  [@Cao2018openpose; @Simon2017hand], which provided the locations of 18 body parts (ears, nose, wrists, etc.). To do so, a convolutional neural network was used for initial anatomical detection, and part affinity fields were subsequently applied for part association to produce a series of body part candidates. Once these body part candidates were matched to a single individual in the frame, they were finally assembled into a pose. Thus, while we only made use of the outputs of the face and hand detections, the entire set of pose information from an individual was used to determine the presence of a face/hand, making the process more robust to occlusion than methods optimized to detect only faces or hands. Note, however, that these face/hand detections are reliant on the detection of at least a partial pose, so some very up-close views of faces/hands may go undetected.

@@

As has been observed in other studies on automated annotation of headcam data [e.g. @frank2013; @sanchez2018postural; @bambach2015lending], detection tasks that are easy in third-person video can be quite challenging in egocentric videos, due to difficult angles and sizes as well as substantial occlusion. For example, the infant perspective often contains non-canonical viewpoints of faces (e.g., looking up at a caregiver's chin) as well as partially-occluded or oblique viewpoints of both faces and hands. Further, hand detection tends to be a harder computational problem than face detection [@Simon2017hand; @bambach2015lending]. We thus expected overall performance to be lower in these naturalistic videos than on either photos taken from the adult perspective or in egocentric videos in controlled, laboratory settings [e.g., @sanchez2018postural].

@@

```{r FacesAndHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces and hands seen as a function of age for each child in the dataset. Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range. Dashed lines show estimated trend lines from proportion of faces/hands in view when detections are restricted to the upper 60 percent of the FOV, reducing the contribution of children's own hands; dotted lines show estimated trend lines for human annotations on the 24K gold set."}

@@

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child.^[All analyses and preprocessed data files for this paper are avaliable at https://tinyurl.com/detecting-social-info] Data from videos were binned according to the age of the child (in weeks). First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:FacesAndHands}); in contrast, we did not observe an increase in the proportion of hands in view, but rather a slight decrease. These effects were quantified with two separate linear mixed-models (see Tables 1 & 2).^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.]

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Figure \ref{fig:FacesAndHands}) or OpenPose annotations on the entire dataset (see solid lines in Figure \ref{fig:FacesAndHands}). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. Nonetheless, one reason this could be the case is the much larger field of view that was captured by the cameras used in this study: These cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible, leading to a larger field of view (109 degrees horizontal x 70 degrees vertical) than in many previous studies; for example, in @Fausey2016 the FOV was 69 x 41 degrees. This larger FOV may have allowed the SAYCam cameras to capture not only the presence of a social partner’s hands interacting with objects or gestures, but also the children’s own hands, leading to more frequent hand detections.

```{r}
# replace NAs with FALSE to compare proportion hands detected
d_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE))
```

As children's hands tended to occur in the lower visual field (see Figure \ref{fig:handLocation}), we thus re-analyzed the entire dataset while restricting our analysis to the upper 60% of the field of view, decreasing the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)*100`% to `r round(mean(d_cropped$hand_detected),2)*100`%, but only decreased face detections from `r round(mean(d_all$face_openpose),2)*100`% to `r round(mean(d_cropped$face_detected),3)*100`%. This cropping likely removed both the majority of detections of children's own hands but also some detections of adult hands (see Figure \ref{fig:handLocation}), especially as OpenPose was biased to miss children's hands when they were in view. Nonetheless, within this modified field of view, we still observed more hand detections than face detections (see dashed lines in Figure \ref{fig:FacesAndHands}).

@@

Here, we analyzed the social information in view in a dense, longitudinal dataset, applying a modern computer-vision model to quantify the proportion of hands and faces seen from each of three children's egocentric perspective from 6 to 32 months of age. This analysis has yielded a better understanding of infants' evolving access to social information. We found a moderate decrease across age in the proportion of faces in view in the videos, in keeping with previous work [@Fausey2016]. This finding is particularly notable given that, in previous cross-sectional data, this effect seems to be most strongly driven by infants younger than 4 months of age [e.g., @Fausey2016; @Jayaraman2015; @Sugden2014] who see both more frequent and more persistent faces [@Jayaraman2018].

@@

A coarse analysis based on the location the videos were filmed in further highlights the variability of the social information in view during different activites, showing differences across locations and between individual children. Within a given, well-defined context---e.g., mealtime in kitchens---S saw more faces than A, and S saw more faces in the kitchen than in other locations. This variability likely stems from the fact that there are at least three ways to feed a young child: 1) sitting in front of the child, facing them as they sit in a high chair; 2) sitting behind the child, holding them as they face outward, and 3) sitting side by side. Each of these positions offer the child differing degrees of visual access to faces and hands. While the social information in view may be variable across children in different activitiy contexts, these analyses suggest they could be stable within a given child's day-to-day experience.
